---
title: "Grid Carbon Intensity by Balancing Authority"
output: html_notebook
---


```{r}
library(tidyverse)
library(httr2)
library(jsonlite)
```

Downloaded data from Electricity Maps
https://www.electricitymaps.com/data-portal/united-states-of-america

Merge balancing authority files
```{r}
temp <- list.files("BA Carbon 2021/", full.names = TRUE)#, pattern = "\\.csv$") 
ba_carbon <- readr::read_csv(temp, id = "file_name")

temp
ba_carbon
```

Register with Electricity Maps - done
Test live carbon intensity call with lat and lon - done
Test get zone in response - done

```{r}

api_key <- Sys.getenv("electricitymaps_key")

lat <- 33.90583	
lon <- -118.40194

req <- request(paste('https://api-access.electricitymaps.com/free-tier/carbon-intensity/latest?','lat=',lat,'&lon=',lon, sep='')) %>% 
  req_headers('auth-token' = api_key) 

#req %>% req_dry_run()

resp <- req_perform(req)
resp_json <- resp %>% resp_body_json()
data_df <- as.data.frame(resp_json)
data_df$zone

```

Create list of unique lat and lon combinations for looking up zones
```{r}
gis_list <- distinct(planned_sites, lat, lon)
```

Write function to input lat/lon and output zone and add zone to "planned_sites"
```{r}
#planned_sites 

get_zones <- function(i) {

#i <- 2
lat_i <- gis_list[i,"lat"]
lon_i <- gis_list[i,"lon"]

req <- request(paste('https://api-access.electricitymaps.com/free-tier/carbon-intensity/latest?','lat=',lat_i,'&lon=',lon_i, sep='')) %>% 
  req_headers('auth-token' = api_key) 

resp <- req_perform(req)
resp_json <- resp %>% resp_body_json()
data_df <- as.data.frame(resp_json)
zone <- data_df$zone

planned_sites <<- planned_sites %>% 
  rows_update(tibble(lat = lat_i, lon = lon_i, grid_zone = zone), by = c("lat","lon"))
  
}


```

Write a function to suppress message output:
```{r}
quiet <- function(x) { 
  sink(tempfile()) 
  on.exit(sink()) 
  invisible(force(x)) 
} 
```

Run get_zones function on all locations
```{r}

for (i in 1:nrow(gis_list)) {
    #print(i)
    #currentTime <- Sys.time()
    #print(currentTime)
    quiet(get_zones(i))
    Sys.sleep(0.1) #so as not to exceed the rate limit for the API
}

planned_sites

```


Match on ba code at end of Zone Id to "ba_carbon"
Identify grid 2021 average for the zone/balancing authority for each site
```{r}
site_carbon <- planned_sites %>% 
  left_join(ba_carbon, by = join_by(grid_zone == 'Zone Id')) %>% 
  rename(co2e_direct = 'Carbon Intensity gCO₂eq/kWh (direct)', co2e_lca = 'Carbon Intensity gCO₂eq/kWh (LCA)')

```


Filter on normalised capacity to get green (electrolized) H2 only
```{r}
green_sites <- 
site_carbon %>% 
  filter(`Normalised capacity (MWel)` > 0 )
```

Site emissions * estimated electricity usage
Calculate green h2 site emissions based on listed renewables LCA from elsewhere
Compare to emissions if totally sourced from grid (above)

Plot green H2 sites with color indicating carbon emissions level

Look at CO2/H2 for gray, green, blue sites
Gray - use data from hydrogen_facilities (weed out green/blue sites?)
Green - use type of renewables LCA reference from elsewhere
Blue - estimate total CO2 byproduct and subtract captured? 
        - Use data from hydrogen facilities or find other reference?
