---
title: "Hydrogen Low-Carbon Projects Data Set"
output: html_notebook
---

Load necessary packages
```{r}
library(tidyverse)
library(googlesheets4)
#install.packages("usmap")
library(usmap) #import the package
#library(ggplot2) #use ggplot2 to add layer for visualization
#install.packages("sp")
#install.packages("sf")
library(sp)
library(sf)

```

Load hydrogen projects data from google sheet
```{r}
gs4_deauth()
hydrogen_projects <- read_sheet("https://docs.google.com/spreadsheets/d/17YCMTTdN7yf_oA8eriyD9tPt1SePWWQmj-AvJY5lbiE/edit#gid=1245935679", sheet = "Projects")

hydrogen_projects <- hydrogen_projects %>% 
  mutate(type = case_when(
          `Normalised capacity (MWel)` > 0 ~ 'green',
          `Normalised capacity (t CO₂ captured/y)` > 0 ~ 'blue')
          )
  
hydrogen_projects 
```

Test plotting lat and lon
```{r}
hydrogen_projects_df <- as.data.frame(hydrogen_projects)

planned_sites <- hydrogen_projects_df %>% 
  rename(lat = Latitude, lon = Longitude) %>% #add other fields here
  drop_na(lat,lon) 

planned_sites_t <- usmap_transform(planned_sites)

plot_usmap(regions = "states") + 
  geom_point(data=planned_sites_t, aes(x=x, y=y, color=type)) +
  ggtitle("Sample of Planned US Green & Blue Hydrogen Production Facilities") +
  theme(legend.position = "right") +
  labs(colour = "H2 Type") +
  scale_color_manual(values = c("blue" = "blue",
                                "green"="green")) 

```

Pull data from GridCarbonIntensity.Rmd

Plot green sites
- Size by grid carbon intensity
- Size by renewable percentage

```{r}
#green_sites

green_sites_t <- usmap_transform(green_sites)

plot_usmap(regions = "states") + 
  geom_point(data=green_sites_t, aes(x=x, y=y, size=co2e_lca), color="brown") 

plot_usmap(regions = "states") + 
  geom_point(data=green_sites_t, aes(x=x, y=y, size=`Renewable Percentage`), color="green")

```


Zone emissions * estimated electricity usage
Calculate green h2 site emissions based on listed renewables LCA from elsewhere
Compare to emissions if totally sourced from grid (above)

```{r}
#green_sites

green_sites %>% 
  mutate(ratio = `Normalised capacity (nm³ H₂/h)`*1.0/`Normalised capacity (kt H2/y)`, el_ratio = `Normalised capacity (MWel)`*1.0/`Normalised capacity (kt H2/y)`)

#ratio of nm³/h to kt/y is a constant 1282.643
#ratio of MW capacity to kt/y is a constant per electrolysis method

#use estimate that 50 kWh electricity required to produce 1 kg of H2
#https://en.wikipedia.org/wiki/Hydrogen_economy
#co2e_lca = 'Carbon Intensity gCO₂eq/kWh (LCA)'

#use estimate of 50 g/kWh for solar LCA
#https://www.nrel.gov/docs/fy13osti/56487.pdf

#g CO2/kWh * 1 t CO2/1,000,000 g CO2 * 50 kWh/kg H2 * 1,000,000 kg H2/ kt H2 * kt H2 / y => t CO2 / y

green_sites <- green_sites %>% 
  mutate(est_elec_GWh_y = 50 * `Normalised capacity (kt H2/y)`) %>% 
  mutate(t_co2e_y_solar = est_elec_GWh_y*50, t_co2e_y_grid = co2e_lca * 50 * `Normalised capacity (kt H2/y)`) %>% 
  mutate(co2_h2_ratio_solar = t_co2e_y_solar/1000/`Normalised capacity (kt H2/y)`, co2_h2_ratio_grid = t_co2e_y_grid/1000/`Normalised capacity (kt H2/y)`, solar_pct = t_co2e_y_solar/t_co2e_y_grid) 

green_sites

#~21 co2:h2 ratio tracks
#https://www.sciencedirect.com/science/article/pii/S2590174520300155

```

```{r}
green_sites_table <- green_sites %>% 
  mutate(solar_co2_reduction_pct = 100*(1-solar_pct)) %>% 
  select(`Project name`, `Company`, `Date online`,`Normalised capacity (kt H2/y)`, grid_zone, solar_co2_reduction_pct, co2_h2_ratio_solar, co2_h2_ratio_grid)

write_csv(green_sites_table, file = "green_sites_table.csv")

```



Plot green H2 sites with color indicating carbon emissions level

```{r}

green_sites_t <- usmap_transform(green_sites)

plot_usmap(regions = "states") + 
  geom_point(data=green_sites_t, aes(x=x, y=y, size=t_co2e_y_grid), color="brown") +
  ggtitle("Life-Cycle Emissions Based on Grid Electricity") +
  theme(legend.position = "right") +
  labs(size = "Metric Tons CO2/Year")  


plot_usmap(regions = "states") + 
  geom_point(data=green_sites_t, aes(x=x, y=y, size=t_co2e_y_solar), color="brown") +
  ggtitle("Life-Cycle Emissions Based on Solar LCA") +
  theme(legend.position = "right") +
  labs(size = "Metric Tons CO2/Year") 
```

Look at CO2/H2 for gray, green, blue sites
https://www.sciencedirect.com/science/article/pii/S2590174520300155
2.5 for green
7 for blue
10-15 for gray
25 for grid-powered electrolysis

Gray - use data from hydrogen_facilities (weed out green/blue sites?)
Green - use type of renewables LCA reference from elsewhere
Blue - estimate total CO2 byproduct and subtract captured? 
        - Use data from hydrogen facilities or find other reference?

